---
title: "Daily Weights Data Cleaning"
hitheme: tomorrow
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
highlighter: highlight.js
---

# Loading Data and Cleaning


```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(here)
library(skimr)
library(lubridate)
library(data.table)
library(zoo)
library(ggpubr)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

# load daily weights
weights <- read_csv(here("Body_Trace_Weights_C1-C3_4.20.20.csv"), col_types = "ccdd") %>% clean_names()

# load participant ids
excel_sheets(here::here("17-0369_Body_Trace_Scale_Assignment.xlsx"))

cohort1 = read_excel(here::here("17-0369_Body_Trace_Scale_Assignment.xlsx"), 
           sheet = "C1") %>%
  clean_names()

cohort1_errors = read_excel(here::here("17-0369_Body_Trace_Scale_Assignment.xlsx"), 
           sheet = "C1 Errors") %>%
  clean_names()

cohort2 = read_excel(here::here("17-0369_Body_Trace_Scale_Assignment.xlsx"), 
           sheet = "C2") %>%
  clean_names()

cohort3 = read_excel(here::here("17-0369_Body_Trace_Scale_Assignment.xlsx"), 
           sheet = "C3") %>%
  clean_names()

cohort1 = cohort1 %>% 
  mutate(cohort = 1) %>%
  select(participant_id, imei, cohort, scale_id)

cohort2 = cohort2 %>% 
  mutate(cohort = 2) %>%
  select(participant_id, imei, cohort)

cohort3 = cohort3 %>% 
  mutate(cohort = 3) %>%
  select(participant_id, imei, cohort)

cohorts <- full_join(cohort1, cohort2) %>% full_join(.,cohort3)
weights = left_join(weights, cohorts)
weights$cohort <- factor(weights$cohort, c("1", "2", "3"))

#removing staff data
weights = weights %>% filter(!participant_id =="DRIFT2 Staff")

# Removing early withdrawal patient 

weights = weights %>% filter(!participant_id =="CWE-124")

# Combining new and old weights for participants 
weights$participant_id = str_sub(weights$participant_id, 1, 7)

#creating weight variable in pounds
weights = weights %>% mutate(wt_lb=wt_kg*2.2046226218)

## Switching the weights from spouses JUT-032 and KCE-034 when they were using each other's scales. 

weights = weights %>% mutate(participant_id=
    case_when(
      participant_id=="JUT-032" & wt_lb > 215 ~ "JUT-032",
      participant_id=="JUT-032" & wt_lb <=215 ~ "KCE-034",
      participant_id=="JUT-032" & wt_lb == NA  ~ "JUT-032",
      !participant_id=="JUT-032" ~ participant_id,
      TRUE ~ participant_id,
        ))

weights = weights %>% mutate(participant_id=
    case_when(
      participant_id=="KCE-034" & wt_lb < 215 ~ "KCE-034",
      participant_id=="KCE-034" & wt_lb >=215 ~ "JUT-032",
      participant_id=="KCE-034" & wt_lb == NA  ~ "KCE-034",
      !participant_id=="KCE-034" ~ participant_id,
      TRUE ~ participant_id,
))

## Combining weights from YOL-103 into YOR-103 

weights = weights %>% mutate(participant_id=
  case_when(
    participant_id== "YOL-103" ~ "YOR-103",
    participant_id!= "YOL-103" ~ participant_id,
    TRUE ~ participant_id,
  ))


weights$date_time_of_wt <- mdy_hm(weights$date_time_of_wt)
#grabing just the date without time
split_dates <- str_split_fixed(weights$date_time_of_wt, " ", 2)
weights$weight_dates <- split_dates[,1] %>% ymd()

#study days variable
weights = weights %>% mutate(weight_date_numeric = as.numeric(weight_dates), weighed_in = "yes") %>% group_by(participant_id)%>% mutate(study_days = (weight_date_numeric - min(weight_date_numeric)) + 1, observation_days = max(study_days)) %>% ungroup()

## Removing weight observations past day 600 

weights = weights %>% mutate(study_days_outliers = if_else(study_days<=600, 1, 0)) %>% filter(study_days_outliers=="1") %>% select(-(study_days_outliers))

## splitting up study days into fourths 
weights <-weights %>% group_by(participant_id) %>% 
  mutate(study_time=(max(weight_dates) -min(weight_dates))) %>% 
  mutate(q1=min(weight_dates)+study_time*(1/4)) %>% 
  mutate(q2=min(weight_dates)+study_time*(2/4)) %>% 
  mutate(q3=min(weight_dates)+study_time*(3/4)) %>% 
  mutate(q4=max(weight_dates)) %>% 
  mutate(period=if_else(weight_dates<q1, 1, if_else(weight_dates>=q1 & weight_dates<q2, 2,
    if_else(weight_dates>=q2 & weight_dates<q3, 3, 
    if_else(weight_dates>=q3 & weight_dates<=q4,4,0))))) %>% ungroup() 
weights <- weights %>% select(imei, date_time_of_wt, wt_kg, tare_kg, 
        participant_id,cohort, scale_id, weight_dates, wt_lb, period, weight_dates, study_days) %>% ungroup()
weights$period <- factor(weights$period, levels = (c("1", "2", "3", "4"))) 

## Adding weight dates variable, weight_dates_join to join with drift sets 

weights$weight_dates_join = weights$weight_dates


## Loading Drift2 Database 

data <- read_csv(here("DRIFT2Database-AllWeightsInC1C3AsOf_DATA_2020-06-09_1016.csv")) %>% clean_names()
data$initials = na.locf(data$initials)

## Modifying record ID then attaching it to initials

data$record_id = str_pad(data$record_id, 3, "left", "0")
data$participant_id = paste(data$initials, data$record_id, sep="-")

data = data %>% mutate(participant_id=
  case_when(
    participant_id== "rVCA-041" ~ "VCA-041",
    participant_id!= "VCA-041" ~ participant_id,
    TRUE ~ participant_id,
  ))

## Grabbing weight values and dates for the classes and merging them into two columns using a gather and melt function. 

data_class_weights <- data %>% select(participant_id, cohort_number, redcap_event_name, class_wt_wk0, class_wt_wk1,class_wt_wk2,class_wt_wk3,class_wt_wk4,class_wt_wk5,class_wt_wk6,class_wt_wk7,class_wt_wk8,class_wt_wk9,class_wt_wk10,class_wt_wk11,class_wt_wk12,class_wt_wk13,class_wt_wk14,class_wt_wk15,class_wt_wk16,class_wt_wk18,class_wt_wk20,class_wt_wk22,class_wt_wk24,class_wt_wk26,class_wt_wk28,class_wt_wk30,class_wt_wk32,class_wt_wk34,class_wt_wk36,class_wt_wk38,class_wt_wk40,class_wt_wk42,class_wt_wk44,class_wt_wk46,class_wt_wk48,class_wt_wk50,class_wt_wk52, class_dt_wk0, class_dt_wk1, class_dt_wk2, class_dt_wk3,class_dt_wk4,class_dt_wk5,class_dt_wk6,class_dt_wk7,class_dt_wk8,class_dt_wk9,class_dt_wk10,class_dt_wk11,class_dt_wk12,class_dt_wk13,class_dt_wk14,class_dt_wk15,class_dt_wk16,class_dt_wk18,class_dt_wk20,class_dt_wk22,class_dt_wk24,class_dt_wk26,class_dt_wk28,class_dt_wk30,class_dt_wk32,class_dt_wk34,class_dt_wk36,class_dt_wk38,class_dt_wk40,class_dt_wk42,class_dt_wk44,class_dt_wk46,class_dt_wk48,class_dt_wk50,class_dt_wk52)

data_class_weights= data_class_weights %>% filter(redcap_event_name=="screening_arm_1")

data_class_weights <- melt(setDT(data_class_weights), measure = patterns("^class_wt_wk", "^class_dt_wk"), value.name = c("weight_lb_class", "weight_dates_class"))

data_class_weights= data_class_weights%>% select(participant_id, cohort_number, weight_lb_class, weight_dates_class)
data_class_weights = as_tibble(data_class_weights)
data_class_weights$weight_dates_class = mdy(data_class_weights$weight_dates_class)
#duplicating dates to join and classify all weights into same period
data_class_weights$weight_dates_join = data_class_weights$weight_dates_class
#duplicating weights into a research weights column to be joined with outcomes 
data_class_weights$research_weights = data_class_weights$weight_lb_class

## Grabbing bluetooth weights and dates and merging them into two columns using a gather and melt function. 

data_class_weights_bt <- data %>% select(participant_id, cohort_number, redcap_event_name, class_wt_wk0_bt, class_wt_wk1_bt,class_wt_wk2_bt,class_wt_wk3_bt,class_wt_wk4_bt,class_wt_wk5_bt,class_wt_wk6_bt,class_wt_wk7_bt,class_wt_wk8_bt,class_wt_wk9_bt,class_wt_wk10_bt,class_wt_wk11_bt,class_wt_wk12_bt,class_wt_wk13_bt,class_wt_wk14_bt,class_wt_wk15_bt,class_wt_wk16_bt,class_wt_wk18_bt,class_wt_wk20_bt,class_wt_wk22_bt,class_wt_wk24_bt,class_wt_wk26_bt,class_wt_wk28_bt,class_wt_wk30_bt,class_wt_wk32_bt,class_wt_wk34_bt,class_wt_wk36_bt,class_wt_wk38_bt,class_wt_wk40_bt,class_wt_wk42_bt,class_wt_wk44_bt,class_wt_wk46_bt,class_wt_wk48_bt,class_wt_wk50_bt,class_wt_wk52_bt, class_dt_wk0_bt, class_dt_wk1_bt, class_dt_wk2_bt, class_dt_wk3_bt,class_dt_wk4_bt,class_dt_wk5_bt,class_dt_wk6_bt,class_dt_wk7_bt,class_dt_wk8_bt,class_dt_wk9_bt,class_dt_wk10_bt,class_dt_wk11_bt,class_dt_wk12_bt,class_dt_wk13_bt,class_dt_wk14_bt,class_dt_wk15_bt,class_dt_wk16_bt,class_dt_wk18_bt,class_dt_wk20_bt,class_dt_wk22_bt,class_dt_wk24_bt,class_dt_wk26_bt,class_dt_wk28_bt,class_dt_wk30_bt,class_dt_wk32_bt,class_dt_wk34_bt,class_dt_wk36_bt,class_dt_wk38_bt,class_dt_wk40_bt,class_dt_wk42_bt,class_dt_wk44_bt,class_dt_wk46_bt,class_dt_wk48_bt,class_dt_wk50_bt,class_dt_wk52_bt)

data_class_weights_bt= data_class_weights_bt %>% filter(redcap_event_name=="screening_arm_1")

data_class_weights_bt <- melt(setDT(data_class_weights_bt), measure = patterns("^class_wt_wk", "^class_dt_wk"), value.name = c("weight_lb_class_bt", "weight_dates_class_bt"))

data_class_weights_bt = data_class_weights_bt %>% select(participant_id, cohort_number, weight_lb_class_bt, weight_dates_class_bt)
data_class_weights_bt = as_tibble(data_class_weights_bt)
data_class_weights_bt$weight_dates_class_bt = mdy(data_class_weights_bt$weight_dates_class_bt)
data_class_weights_bt$weight_dates_join = data_class_weights_bt$weight_dates_class_bt


## data set for the withdraw dates 

data_withdraw = data %>% filter(redcap_event_name=="screening_arm_1") %>%  select(participant_id, withdrawal_date, withdrawal_wt_lbs, withdrawal_wt_date )
data_withdraw$withdrawal_wt_date = mdy(data_withdraw$withdrawal_wt_date)
data_withdraw$withdrawal_date = mdy(data_withdraw$withdrawal_date)
data_withdraw = data_withdraw %>% group_by(participant_id) %>% mutate(participant_id_numbered=paste(participant_id, (seq(1:length(participant_id)))+2000)) %>% ungroup()
data_withdraw$weight_dates_join = data_withdraw$withdrawal_date
data_withdraw$weight_dates = data_withdraw$withdrawal_date
## The reason for adding the withdrawal dates into the weight_dates column is so that during analysis, when trying to find the number of weigh-in observations per month, the proper denominator can be used. Leaving in withdrawal participants in the initial months and removing them from the denominator after having withdrawn. 

## data set for the outcome weights and outcome weight dates 

data_weight_outcomes = data %>%  select(participant_id,record_id, outcome_wt_date, outcome_avg_wt_lbs)
data_weight_outcomes$outcome_wt_date = mdy(data_weight_outcomes$outcome_wt_date)
data_weight_outcomes = data_weight_outcomes %>% group_by(record_id) %>% mutate(participant_id_numbered=paste(participant_id, (seq(1:length(participant_id)))+3000)) %>% ungroup()
#duplicating weights to research weights to be combined with class weights 
data_weight_outcomes$research_weights = data_weight_outcomes$outcome_avg_wt_lbs

#correcting participant EJO-092 outcome weight date: improperly put in as 2014 instead of 2019 

data_weight_outcomes = data_weight_outcomes %>% mutate(outcome_wt_date = 
        case_when(
          outcome_wt_date == "2014-10-22" ~ as.Date("2019-10-22"),
          TRUE ~ outcome_wt_date,
        ))


data_weight_outcomes$weight_dates_join = data_weight_outcomes$outcome_wt_date


## creating a new variable, participant_id_numbered, to attempt to join data sets without duplicates

weights = weights %>% group_by(participant_id) %>% mutate(participant_id_numbered=paste(participant_id, seq(1:length(participant_id)))) %>% ungroup()

data_class_weights = data_class_weights %>% group_by(participant_id) %>% mutate(participant_id_numbered=paste(participant_id, (seq(1:length(participant_id)))+500)) %>% ungroup()

data_class_weights_bt = data_class_weights_bt %>% group_by(participant_id) %>% mutate(participant_id_numbered=paste(participant_id, (seq(1:length(participant_id)))+1000)) %>% ungroup()

## Joining the blue tooth weights, the class weights, and the daily weights data. 

daily_drift_data = full_join(weights, data_class_weights_bt) %>% full_join(.,data_class_weights) %>% full_join(.,data_withdraw) %>% full_join(.,data_weight_outcomes)
daily_drift_data = daily_drift_data %>% select(-participant_id_numbered)

## Filling in the missing cohort values that came from joining data frames without that variable

daily_drift_data = daily_drift_data %>% group_by(participant_id) %>% fill(cohort) %>% ungroup()
```

## Creating a withdrawal variable 

```{r}
daily_drift_data = daily_drift_data %>% group_by(participant_id) %>% mutate(withdrawal =
        case_when(
          (mean(withdrawal_date, na.rm=TRUE)) > as.Date(01-01-2000) ~1,
           TRUE~0,
          )) %>% ungroup()

#could remove outliers from one data set, but for now will just have to use filter(withdrawal=="0") in any line of code where no withdrawal participants should be there 
#daily_drift_data_rmo = daily_drift_data_rmo %>% filter(withdrawal=="0")

#practice = practice %>% group_by(participant_id) %>%  mutate(study_length = max(withdrawal_date, na.rm=TRUE) - min(weight_dates, na.rm=TRUE)) %>% mutate(withdrawal_removal =      case_when(
#        study_length > 0 & study_length < 200 ~ 1,
#        TRUE ~0
#      ))
#practice = practice %>% filter(withdrawal_removal=="0")
 
```                                                                                     

## 30 day period of study days (weights from the bluetooth scales)
```{r}
daily_drift_data = daily_drift_data %>% group_by(participant_id) %>% 
  mutate(m1=as.numeric(min(weight_dates, na.rm = TRUE) + 29)) %>% 
  mutate(m2=as.numeric(min(weight_dates, na.rm = TRUE) + 59)) %>% 
  mutate(m3=as.numeric(min(weight_dates, na.rm = TRUE) + 89)) %>%  
  mutate(m4=as.numeric(min(weight_dates, na.rm = TRUE) + 119)) %>% 
  mutate(m5=as.numeric(min(weight_dates, na.rm = TRUE) + 149)) %>% 
  mutate(m6=as.numeric(min(weight_dates, na.rm = TRUE) + 179)) %>% 
  mutate(m7=as.numeric(min(weight_dates, na.rm = TRUE) + 209)) %>% 
  mutate(m8=as.numeric(min(weight_dates, na.rm = TRUE) + 239)) %>% 
  mutate(m9=as.numeric(min(weight_dates, na.rm = TRUE) + 269)) %>% 
  mutate(m10=as.numeric(min(weight_dates, na.rm = TRUE) + 299)) %>% 
  mutate(mrem=as.numeric(min(weight_dates, na.rm = TRUE) + 10000)) %>% 
  ungroup()

## Creating a new variable putting dates into 30 day periods in both weight dates (daily weights data) and class dates (redcap data)


daily_drift_data = daily_drift_data %>% group_by(participant_id) %>% mutate(period_30=
  case_when(
    as.numeric(weight_dates) <=m1 ~ 1,
    as.numeric(weight_dates)<=m2  ~ 2,
    as.numeric(weight_dates)<=m3  ~ 3,
    as.numeric(weight_dates)<=m4  ~ 4,
    as.numeric(weight_dates)<=m5 ~ 5,
    as.numeric(weight_dates)<=m6 ~ 6,
    as.numeric(weight_dates)<=m7  ~ 7,
    as.numeric(weight_dates)<=m8  ~ 8,
    as.numeric(weight_dates)<=m9 ~ 9,
    as.numeric(weight_dates)<=m10  ~ 10,
    as.numeric(weight_dates)<=mrem  ~ 11, 
  )) %>%  ungroup()

```

## Removing unneeded variables from data set 
```{r}
daily_drift_data = daily_drift_data %>% select("imei", "participant_id", "weight_dates", "weight_dates_class_bt", "weight_dates_class", "withdrawal_wt_date", "withdrawal_date", "outcome_wt_date", "cohort", "period", "wt_kg", "wt_lb", "study_days", "weight_lb_class_bt", "weight_lb_class", "withdrawal_wt_lbs", "outcome_avg_wt_lbs", "weight_dates_join", "research_weights", "withdrawal", "period_30", "m1", "m2", "m3", "m4", "m5", "m6", "m7", "m8", "m9", "m10", "mrem")
```

```{r}
## Data Manipulation with cohorts 1 and 2 

data_cohorts_1_2 = daily_drift_data %>% filter(cohort %in% c("1", "2"))


```


## Periods for outliers: 15 day (8 periods) / 30 day


```{r}

data_cohorts_1_2 = data_cohorts_1_2 %>% group_by(participant_id) %>% 
  mutate(p1=as.numeric(min(weight_dates_join, na.rm = TRUE) + 14)) %>% 
  mutate(p2=as.numeric(min(weight_dates_join, na.rm = TRUE) + 29)) %>% 
  mutate(p3=as.numeric(min(weight_dates_join, na.rm = TRUE) + 44)) %>%  
  mutate(p4=as.numeric(min(weight_dates_join, na.rm = TRUE) + 59)) %>% 
  mutate(p5=as.numeric(min(weight_dates_join, na.rm = TRUE) + 74)) %>% 
  mutate(p6=as.numeric(min(weight_dates_join, na.rm = TRUE) + 89)) %>% 
  mutate(p7=as.numeric(min(weight_dates_join, na.rm = TRUE) + 104)) %>% 
  mutate(p8=as.numeric(min(weight_dates_join, na.rm = TRUE) + 119)) %>% 
  mutate(p9=as.numeric(min(weight_dates_join, na.rm = TRUE) + 149)) %>% 
  mutate(p10=as.numeric(min(weight_dates_join, na.rm = TRUE) + 179)) %>% 
  mutate(p11=as.numeric(min(weight_dates_join, na.rm = TRUE) + 209)) %>% 
  mutate(p12=as.numeric(min(weight_dates_join, na.rm = TRUE) + 239)) %>% 
  mutate(p13=as.numeric(min(weight_dates_join, na.rm = TRUE) + 269)) %>% 
  mutate(p14=as.numeric(min(weight_dates_join, na.rm = TRUE) + 299)) %>% 
  mutate(p15=as.numeric(min(weight_dates_join, na.rm = TRUE) + 329)) %>% 
  mutate(p16=as.numeric(min(weight_dates_join, na.rm = TRUE) + 359)) %>% 
  mutate(prem=as.numeric(min(weight_dates_join, na.rm = TRUE) + 1000)) %>%
   ungroup()

data_cohorts_1_2 = data_cohorts_1_2 %>% group_by(participant_id) %>% mutate(outlier_periods =
  case_when(
    as.numeric(weight_dates_join) <=p1 ~ 1,
    as.numeric(weight_dates_join)<=p2  ~ 2,
    as.numeric(weight_dates_join)<=p3  ~ 3,
    as.numeric(weight_dates_join)<=p4  ~ 4,
    as.numeric(weight_dates_join)<=p5 ~ 5,
    as.numeric(weight_dates_join)<=p6 ~ 6,
    as.numeric(weight_dates_join)<=p7  ~ 7,
    as.numeric(weight_dates_join)<=p8  ~ 8,
    as.numeric(weight_dates_join)<=p9 ~ 9,
    as.numeric(weight_dates_join)<=p10  ~ 10,
    as.numeric(weight_dates_join)<=p11  ~ 11, 
    as.numeric(weight_dates_join)<=p12  ~ 12,
    as.numeric(weight_dates_join)<=p13  ~ 13,
    as.numeric(weight_dates_join)<=p14 ~ 14,
    as.numeric(weight_dates_join)<=p15  ~ 15,
    as.numeric(weight_dates_join)<=p16  ~ 16,
    as.numeric(weight_dates_join)<=prem  ~ 17,
  )) %>%  ungroup()


```



## Removing same day observations 
```{r}
## removing based on outlier periods 
data_cohorts_1_2 = data_cohorts_1_2 %>% group_by(participant_id, outlier_periods) %>% mutate(mean_research = mean(research_weights, na.rm=TRUE)) %>% ungroup() %>% group_by(participant_id) %>% mutate(mean_research = coalesce(mean_research, NA_real_)) %>% fill(mean_research, .direction="down") %>% ungroup()

data_cohorts_1_2 = data_cohorts_1_2[!is.na(data_cohorts_1_2$outlier_periods), ]

data_cohorts_1_2 = data_cohorts_1_2 %>% group_by(participant_id, study_days) %>% mutate( study_day_single=
  case_when(
    (abs(wt_lb-mean_research)) == min(abs(wt_lb-mean_research)) ~ 1,
    (abs(wt_lb-mean_research)) != min(abs(wt_lb-mean_research)) ~ 0,
    TRUE~1,
    )) %>% ungroup()
data_cohorts_1_2 = data_cohorts_1_2 %>% filter(study_day_single=="1")


data_cohorts_1_2 = data_cohorts_1_2 %>% group_by(participant_id, study_days) %>% distinct() %>% ungroup()
```


## 5% for first 8 periods and then 10% for remainder 
```{r}

data_cohorts_1_2 = data_cohorts_1_2 %>% group_by(participant_id, outlier_periods) %>%  mutate(outliers = case_when(
  outlier_periods <= 8 & (wt_lb>mean_research*1.05 | wt_lb<mean_research*.95) ~1,
  outlier_periods <= 8 & (wt_lb<=mean_research*1.05 | wt_lb>=mean_research*.95) ~0,
  outlier_periods > 8 & (wt_lb>mean_research*1.1 | wt_lb<mean_research*.9) ~1,
  outlier_periods >8 & (wt_lb<=mean_research*1.1 | wt_lb>=mean_research*.9) ~0,
  TRUE ~ 0
)) %>% ungroup() %>% filter(outliers=="0")

```




## Removing outliers from cohort3 based on the outcome dates 
```{r}

data_cohort3 = daily_drift_data %>% filter(cohort=="3")

# Making outlier  period for daily weights and class weight dates

data_cohort3 = data_cohort3 %>% group_by(participant_id) %>% 
  mutate(p1=as.numeric(min(weight_dates_join, na.rm = TRUE) + 14)) %>% 
  mutate(p2=as.numeric(min(weight_dates_join, na.rm = TRUE) + 29)) %>% 
  mutate(p3=as.numeric(min(weight_dates_join, na.rm = TRUE) + 44)) %>%  
  mutate(p4=as.numeric(min(weight_dates_join, na.rm = TRUE) + 59)) %>% 
  mutate(p5=as.numeric(min(weight_dates_join, na.rm = TRUE) + 74)) %>% 
  mutate(p6=as.numeric(min(weight_dates_join, na.rm = TRUE) + 89)) %>% 
  mutate(p7=as.numeric(min(weight_dates_join, na.rm = TRUE) + 104)) %>% 
  mutate(p8=as.numeric(min(weight_dates_join, na.rm = TRUE) + 130)) %>% 
  mutate(prem=as.numeric(min(weight_dates_join, na.rm = TRUE) + 1000)) %>%
  ungroup()

## Creating a new variable putting dates into 30 day periods in both weight dates (daily weights data) and class dates (redcap data)


data_cohort3 = data_cohort3 %>% group_by(participant_id) %>% mutate(outlier_periods=
  case_when(
    as.numeric(weight_dates_join) <=p1 ~ 1,
    as.numeric(weight_dates_join)<=p2  ~ 2,
    as.numeric(weight_dates_join)<=p3  ~ 3,
    as.numeric(weight_dates_join)<=p4  ~ 4,
    as.numeric(weight_dates_join)<=p5 ~ 5,
    as.numeric(weight_dates_join)<=p6 ~ 6,
    as.numeric(weight_dates_join)<=p7  ~ 7,
    as.numeric(weight_dates_join)<=p8  ~ 8,
    as.numeric(weight_dates_join)<=prem  ~ 9, 
  )) %>%  ungroup()
```




# Removing same day observations 
```{r}
data_cohort3 = data_cohort3 %>% group_by(participant_id, outlier_periods) %>% mutate(mean_research = mean(research_weights, na.rm=TRUE)) %>% ungroup() %>% group_by(participant_id) %>% mutate(mean_research = coalesce(mean_research, NA_real_)) %>% fill(mean_research, .direction="down") %>% ungroup()

data_cohort3 = data_cohort3[!is.na(data_cohort3$outlier_periods), ]

data_cohort3 = data_cohort3 %>% group_by(participant_id, study_days) %>% mutate( study_day_single=
  case_when(
    (abs(wt_lb-mean_research)) == min(abs(wt_lb-mean_research)) ~ 1,
    (abs(wt_lb-mean_research)) != min(abs(wt_lb-mean_research)) ~ 0,
    TRUE~1,
    )) %>% ungroup()
data_cohort3 = data_cohort3%>% filter(study_day_single=="1")


data_cohort3 = data_cohort3 %>% group_by(participant_id, study_days) %>% distinct() %>% ungroup()
```



## 5% for first 7 periods and then 10% for remainder 
```{r}

data_cohort3 = data_cohort3 %>% group_by(participant_id, outlier_periods) %>%  mutate(outliers = case_when(
  outlier_periods <= 7 & (wt_lb>mean_research*1.05 | wt_lb<mean_research*.95) ~1,
  outlier_periods <= 7 & (wt_lb<=mean_research*1.05 | wt_lb>=mean_research*.95) ~0,
  outlier_periods > 7 & (wt_lb>mean_research*1.1 | wt_lb<mean_research*.9) ~1,
  outlier_periods > 7  & (wt_lb<=mean_research*1.1 | wt_lb>=mean_research*.9) ~0,
  TRUE ~ 0
)) %>% ungroup() %>% filter(outliers=="0")

```

#Rejoining cohorts 1&2 with cohort 3 after removing outliers 
```{r}
daily_drift_data_cleaned = full_join(data_cohorts_1_2, data_cohort3)

```

```{r}


#removing outliers from AKI-011 based on researcher verification of inaccurate weights
#aki-011 done
daily_drift_data_cleaned = daily_drift_data_cleaned %>% mutate(outliers = case_when(
  participant_id == "AKI-011" & weight_dates== "2019-06-11" ~ 1,
  participant_id == "AKI-011" & weight_dates== "2019-01-26" ~ 1,
  participant_id == "AKI-011" & weight_dates== "2019-04-20" ~ 1,
  participant_id == "AKI-011" & weight_dates== "2019-09-09" ~ 1,
  participant_id=="AKI-011" & weight_dates!= c("2019-06-11", "2019-01-26", "2019-04-20" ,"2019-09-09")  ~0,
  participant_id!="AKI-011" ~ 0,
  TRUE ~ 0,
)) %>% filter(outliers=="0")

##done
daily_drift_data_cleaned = daily_drift_data_cleaned %>% mutate(outliers = case_when(
  participant_id=="LBU-015" & (study_days > 400 & wt_lb > 182) ~ 1,
  participant_id=="LBU-015" & (study_days > 400 & wt_lb <= 182) ~ 0,
  participant_id=="LBU-015" & (study_days <= 400) ~ 0,
  participant_id!="LBU-015" ~ 0,
  TRUE ~ 0,
)) %>% filter(outliers=="0")


## removed nel-04 removals 
daily_drift_data_cleaned = daily_drift_data_cleaned %>% mutate(outliers = case_when(
  participant_id=="NEL-094" & (wt_lb <240) ~ 1,
  participant_id=="NEL-094" &  (wt_lb >=240) ~ 0,
  participant_id!="NEL-094" ~ 0,
  TRUE ~ 0,
)) %>% filter(outliers=="0")

daily_drift_data_cleaned = daily_drift_data_cleaned %>% mutate(outliers = case_when(
  participant_id=="PBE-123" & (wt_lb <185) ~ 1,
  participant_id=="PBE-123" &  (wt_lb >= 185) ~ 0,
  participant_id!="PBE123" ~ 0,
  TRUE ~ 0,
)) %>% filter(outliers=="0")

```


# Removing withdrawal participants
```{r}
daily_drift_no_withdrawal = daily_drift_data_cleaned %>% filter(withdrawal=="0")

```

# adding in additional variables for Samantha

```{R}
confounding = data %>% filter(redcap_event_name=="screening_arm_1") %>%  select(participant_id, age, sex, race, race_other, ethnicity, marital_status, ever_pregnant, pregnant_number, last_pregnancy)

daily_weights_for_samantha = left_join(daily_drift_no_withdrawal, confounding)

```




